---
title: "Practical Machine Learning Coursera"
author: "Dan Kjeldstrøm Hansen"
date: "March 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is the Coursera Practical Machine Learning course assignment.

We are faced with a superviced learning, classification task.
Given the data collected with wearables from 5 different test persons during different
excerzises, and evaluated by a professional instructor. Each excesize is given a classification
A to E according to the accuracy of the ecersize performed.
The task is to see if it is possible to predict the classe (correctness of excesize) given the
various data collected.

The data is provided in two files, one for training and one for testing and can be found here:

http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

We start off by loading the necessary libraries

```{r libraries}
library(caret)
library(randomForest)
library(xgboost)
library(ggplot2)
library(lattice)
```
As the processing is quite CPU-heavy we'll use parallel processing with the use of the
parallel library, which runs on both LINUX and Windows.

```{r parallel}
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)
```
And the we load the datasets

```{r datasets}
testing <- read.csv('pml-testing.csv',na.strings = c('','NA','#DIV/0!'))
training <- read.csv('pml-training.csv',na.strings = c('','NA','#DIV/0!'))
```
Cleaning out the dataset for use in our algorithms

```{r clean}
#removing variables with close to zero variance (using caret package)
#these variables will have close to no influence on the prediction anyway
nzv_train <- nearZeroVar(training,saveMetrics = T)
nzv_test <- nearZeroVar(testing,saveMetrics = T)

myTraining <- training[,nzv_train$nzv == F]
myTesting <- testing[,nzv_test$nzv == F]

#only keep columns which do not only holds NA's
#as they will have no influence on the prediction anyway
tmp_train_names <- colnames(myTraining)[colSums(is.na(myTraining))==0]
tmp_test_names <- colnames(myTesting)[colSums(is.na(myTesting))==0]

myTraining <- myTraining[,tmp_train_names]
myTesting <- myTesting[,tmp_test_names]

#removing timestamps, X and usernames - not relevant for this excerzise
myTraining <- myTraining[,-c(1:7)]
myTesting <- myTesting[,-c(1:7)]

```
To be able to replicate the resuls we set the seed:

```{r seed}
set.seed(223344)
```

And now to the modelling
First we configure the trainControl object:

```{r trainctrlcfg}
tc <- trainControl(method = 'cv', number = 10, allowParallel = T)
```
And then we try 4 different models within the caret package
We'll let caret take care of the cross-validation and preprocessing
each chunk includes the model, the accuracy and a confusionmatrix

```{r models, eval=FALSE, include=TRUE}
# Extended Gradient Boost - linear
mod_xgb <- train(classe ~.,data=myTraining, method='xgbLinear', trControl= tc)
accu_xgb <- round(max(mod_xgb$results$Accuracy)*100,2)
cfm_xgb <- confusionMatrix(mod_xgb)

# Extended Gradient Boost - linear - Preproc
mod_xgb_pp <- train(classe ~.,data=myTraining, method='xgbLinear', trControl= tc,
                 preProcess = c('center','scale'))
accu_xgb_pp <- round(max(mod_xgb_pp$results$Accuracy)*100,2)
cfm_xgb_pp <- confusionMatrix(mod_xgb_pp)

# Random Forest - out-of-the-box
mod_rf <- train(classe ~.,data=myTraining,method = 'rf')
accu_rf <- round(max(mod_rf$results$Accuracy)*100,2)
cfm_rf <- confusionMatrix(mod_rf)

# Random Forest with both preprocessing and cross-validation  
mod_rf_all <- train(classe ~.,data=myTraining,method = 'rf',
                   trControl = tc,
                   preProcess=c('center','scale'))
accu_rf_all <- round(max(mod_rf_all$results$Accuracy)*100,2)
cfm_rf_all <- confusionMatrix(mod_rf_all)

# Return to single processing
stopCluster(cluster)
registerDoSEQ()
```
Listing the accuracies for comparison:

```{r accuracies}
accuracies <- rbind(accu_xgb,accu_xgb_pp,accu_rf,accu_rf_all)
print(accuracies)
```
Listing the confusion matrices:
```{r}
print(round(cfm_xgb$table,2))
print(round(cfm_xgb_pp$table,2))
print(round(cfm_rf$table,2))
print(round(cfm_rf_all$table,2))
```



Making the prediction for the quiz
```{r predictions}
pred_xgb <- predict(mod_xgb,myTesting)
pred_xgb_pp <- predict(mod_xgb_pp,myTesting)
pred_rf <- predict(mod_rf,myTesting)
pred_rf_all <- predict(mod_rf_all,myTesting)
```

Listing the predictions for comparison
```{r list_predictions}
# All 4 predictions results in the same output, hence using this for validation
# in the quiz, and resulting in a 100% score
preds <- rbind(as.character(pred_xgb),as.character(pred_xgb_pp),
               as.character(pred_rf),as.character(pred_rf_all))
print(preds)
```

This concludes the assignment